---
title: "Project 1 R Markdown"
author: "Ye-in Jeon, Sam Kupfer, Amy Mykityshyn, Krystin Sinclair, Vishnupriya Venkateswaran"
date: "February 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Load data and drop columns with missing data
```{r loaddata}
oecd <- data.frame(read.csv('~/Downloads/OECD_151617.csv'))
oecd <- oecd[sapply(oecd, function(x) !any(is.na(x)))]
head(oecd)
```


### Structure of Dataframe
```{r struct}
str(oecd)
```


### Descriptive Statistics
summary function is used to get mean, min, max, and quartiles for each column
```{r summ}
summary(oecd)
```
And this code to get the standard deviations
```{r sds}
apply(oecd, 2, sd)
```


### Life Satisfaction Geomap
You need the rworldmap package for this part
```{r geomap}
# install.packages("rworldmap")
library(rworldmap)
oecd17 <- subset(oecd, subset = Year == 2017)
oecd17_joined <- joinCountryData2Map(oecd17, joinCode ="NAME", nameJoinColumn = "Country")
mapCountryData(oecd17_joined, nameColumnToPlot="Life.satisfaction", colourPalette = c("red", "yellow", "green"), missingCountryCol = "gray89", addLegend = TRUE, mapTitle = "Life Satisfaction in OECD Countries" )
```


### Data Visualization
```{r data viz 2}
for (col in 2:ncol(oecd)) {
  if (names(oecd)[col] != "Year") {
    nicename <- gsub("\\.", " ", names(oecd)[col])
    hist(oecd[,col], xlab=nicename, main=nicename, col="lightgreen")
    boxplot(oecd[,col], main=nicename, col="darkgoldenrod1")
  }
}

```


### Correlations with Life Satisfaction and Correlation Plot
The corrplot package is needed for this
```{r corrplots, fig.height = 15, fig.width = 20}
# install.packages("corrplot")
library(corrplot)
oecdused <- names(oecd)[names(oecd) != "Year" & names(oecd) != "Country"]
columnsubset <- oecd[oecdused]
oecd_cor <- cor(columnsubset, method = "pearson")
lifesat_cors <- oecd_cor[1,]
lifesat_cors
corrplot(oecd_cor, method = "number", type = "lower")
```
